From Reversible Programs to \\ Univalent Universes and Back

### 2021-07-09 Pi+/paper/equivalence.tex
Rest of styling

To go back from $\PiHatLang$ to $\PiPlusLang$, we turn a natural number into a $\PiPlusLang$ type, using right-biased
addition, that is, the natural number $n$ gets mapped to the type $\onet + (\onet + (\onet + \ldots + \zerot))$. Since
the types are already right-biased, an adjacent transposition in $\PiHatLang$ is easily encoded by using the braiding in
$\PiPlusLang$.

\begin{figure}
  \[
     \Tree [ {\tiny I} [ {\tiny I} {\tiny n} ] ] ~\xrightarrow{\assoclp}~
     \Tree [ [ {\tiny I} {\tiny I} ] {\tiny n} ] ~\xrightarrow[\phantom{xx}\swapp]{\idc\phantom{xx}}~
     \Tree [ [ {\tiny I} {\tiny I} ] {\tiny n} ] ~\xrightarrow{\assocrp}~
     \Tree [ {\tiny I} [ {\tiny I} {\tiny n} ] ] ~
   \]
 \label{fig:plusplusswap}
 \caption{Transpositions from braiding}
\end{figure}

@@

Finally, we show how to interpret $\PiHatLang$ to $\UFin$, and go back from $\UFin$ to $\PiHatLang$.

Types in $\PiHatLang$ are interpreted as 0-cells in $\UFin$, that is, a natural number $n$ is mapped to $\Fin[n]$. The
1-combinators in $\PiHatLang$ are mapped to 1-paths in $\UFin$, that is, 1-loops $\Aut[\Fin[n]]$, in each connected
component. In $\PiHatLang$, the 1-combinators are generated by adjacent transpositions, so these can be easily mapped to
words in $\Sn$ and then to automorphisms using~\Cref{prop:sn-lehmer-fin-equiv}. Finally, 2-combinators are mapped to
2-paths between loops in $\UFin$.

@@

0-cells in $\UFin$ are mapped to their cardinalities in $\PiHatLang$, 1-loops are decoded to words in $\Sn$ to generate
a sequence of adjacent transpositions, producing a 1-combinator in $\PiHatLang$. Finally, 2-paths are quoted back to
2-combinators in $\PiHatLang$.

@@

$\UFin$ as a 1-groupoid, is a (locally strict) weak $(2,0)$-category, and we use this to state our final equivalence.

>>>>

To go back from $\PiHatLang$ to $\PiPlusLang$, we turn a natural number into a $\PiPlusLang$ type, using right-biased
addition, that is, the natural number $n$ gets mapped to the type $\onet + (\onet + (\onet + \ldots + \zerot))$. Since
the types are already right-biased, an adjacent transposition in $\PiHatLang$ is easily encoded by using the braiding in
$\PiPlusLang$, as shown in~\cref{fig:transpfrombraid}. Again, these are shown to be coherent.

\begin{figure}
  \[
     \Tree [ {\tiny I} [ {\tiny I} {\tiny n} ] ] ~\xrightarrow{\assoclp}~
     \Tree [ [ {\tiny I} {\tiny I} ] {\tiny n} ] ~\xrightarrow[\swapp\phantom{xx}]{\phantom{xx}\idc}~
     \Tree [ [ {\tiny I} {\tiny I} ] {\tiny n} ] ~\xrightarrow{\assocrp}~
     \Tree [ {\tiny I} [ {\tiny I} {\tiny n} ] ] ~
   \]
 \label{fig:transpfrombraid}
 \caption{Transpositions from braiding}
\end{figure}

@@

Finally, we show how to interpret $\PiHatLang$ to $\UFin$, and back from $\UFin$ to $\PiHatLang$. Types in $\PiHatLang$
are interpreted as 0-cells in $\UFin$, that is, a natural number $n$ is mapped to $\Fin[n]$. The 1-combinators in
$\PiHatLang$ are mapped to 1-paths in $\UFin$, that is, 1-loops $\Aut[\Fin[n]]$, in each connected component. In
$\PiHatLang$, the 1-combinators are generated by adjacent transpositions, so these can be mapped to words in $\Sn$ and
then to automorphisms using~\Cref{prop:sn-lehmer-fin-equiv}. Finally, 2-combinators are mapped to 2-paths between loops
in $\UFin$.

@@

\noindent 0-cells in $\UFin$ are mapped to their cardinalities in $\PiHatLang$, 1-loops are decoded to words in $\Sn$ to
generate a sequence of adjacent transpositions, producing a 1-combinator in $\PiHatLang$. Finally, 2-paths are quoted
back to 2-combinators in $\PiHatLang$.

@@

% $\UFin$ as a 1-groupoid, is a (locally strict) weak $(2,0)$-category, and we use this to state our final equivalence.
